---
title: "Class 12"
author: "Juan Gonzalez (PID: A69036681)"
format: pdf
editor: visual
---

## 

#Section 1: Proportion og G/G in a population

```{r}
mxl<- read.csv("1000_Genomes_Project_Phase_3-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

#head(mxl)
```

```{r}
#table(mxl$Genotype...forward.strand.)/nrow(mxl)
```

#Section 4: Population Scale Analysis

**Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes**

```{r}
expr<-read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```
```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
colnames(expression)
```


```{r}
median_expression <- tapply(expr$exp, expr$geno, median)
print(median_expression)
```

**Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?**

```{r}
library(ggplot2)
```

```{r}
ggplot(expr)+aes(geno, exp, fill=geno)+
  geom_boxplot(notch=TRUE)
```
G/G is expressed down compared to A/A. The SNP will affect ORMDL3 because of the reduced expression of G/G. 
